---
title: 'Assignment 04: Data Preparation'
author: "Cyril Kombem"
date: "7/28/2021"
output: html_document
---


To open this R markdown file, I first downloaded the Rmarkdown package using install.packages(“rmarkdown") and pulled it into my library using library(rmarkdown). package

#Installing Rmarkdown on R studio
```{r}
install.packages("rmarkdown", repos = "http://cran.us.r-project.org")
```

```{r}
library(rmarkdown)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment= NA)
```


#Discuss the business problem/goal.

In this R Project the researcher wishes to use machine learning techniques to build a recommendation engine that recommends movies to users.
This R project is designed to help you understand the functioning of how a recommendation system works.


#Identify where the dataset was retrieved from.

In order to build our recommendation system, we have used the MovieLens Dataset.
The movies.csv and ratings.csv file contains the data used for this project. This data consists of 105,339 ratings applied over 10,329 movies.



#Loading useful packages for the project.

```{r}
install.packages("recommenderlab", repos = "http://cran.us.r-project.org")
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
install.packages("reshape2", repos = "http://cran.us.r-project.org")
```


```{r}
library(recommenderlab)
library(ggplot2)                       
library(data.table)
library(reshape2)
```

#Identify the code that imported and saved your dataset in R.

1. Importing the csv file into RStudio global environment from GitHub.
In order to read the file into R, 

2. I copied and pasted the raw text file link from GitHub and insert it into the R code using the function below and read the data into R as `drinksdata`

```{r}
download.file(url = "https://raw.githubusercontent.com/ckombem/Assignment-04---Data-Preparation/main/movies.csv" , destfile = "movies.csv")
movie_data <-read.csv("movies.csv", stringsAsFactors =FALSE)
```

```{r}
download.file(url = "https://raw.githubusercontent.com/ckombem/Assignment-04---Data-Preparation/main/ratings.csv" , destfile = "ratings.csv")
rating_data <-read.csv("ratings.csv", stringsAsFactors =FALSE)
```


#Describe your data set (using the common attributes such as #rows, #columns, variable names,
types, means, SD, min/max, etc...).

```{r}
summary(movie_data)
summary(rating_data)
```
1. The variable names are movie_data for the movies and rating_data for ratings.
2. The movies dataset has `r ncol(movie_data)` columns and `r nrow(movie_data)` rows.
3. The rating dataset has `r ncol(rating_data)` columns and `r nrow(rating_data)` rows.
4. The summary details are displayed below.

```{r}
head(movie_data)
head(rating_data)
```

#Discuss any data preparation and errors.

1. Genres in the movies data set is not in a useful format. We will create a one-hot encoding to create a matrix that comprises of corresponding genres for each of the films. 

2. We will remove first row, which was the genre list. Then we will convert the data from characters in to useful integers.

```{r}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors=FALSE)
library(data.table)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', 
                                   type.convert=TRUE), 
                         stringsAsFactors=FALSE) 
colnames(movie_genre2) <- c(1:10)

list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre

for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index,col]) 
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
str(genre_mat2)
```


3. After we we have specified that 1 = Yes and 0 = No. The code below ensure that titles are matched with respective indicators.Below we can see that Heat is an Action Movie and a Crime Movie. Where as Jumanji is an Adventure and it is for Children.


```{r}
SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
head(SearchMatrix)
```

4. For our movie recommendation system to make sense of our ratings through recommenderlabs, we have to convert our matrix into a sparse matrix one. This new matrix is of the class ‘realRatingMatrix’.

```{r}
ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1]) #remove userIds
```

5. Convert rating matrix into a recommenderlab sparse matrix for data to make sense. This new matrix class will be ‘realRatingMatrix’.

```{r}
ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix
```
6.At this point we know that there are 105,339 matrix ratings in this model.

#Discuss the modeling.

This model is designed to use the data type and their respective ratings to recommend movies to interested parties. If the model detects that a user has watched a movie which is action, adventure and animation, it will recommend a movie with other movies having at least one similarity based on the matrix.


```{r}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```  

```{r}
lapply(recommendation_model, "[[", "description")
```
We are implementing an "Item Based Collaborative Filtering". 

This will help apply the formula on the recommendation model and bring back the desired description.

```{r}
recommendation_model$IBCF_realRatingMatrix$parameters
```


#Produce and discuss the output.

Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users. For example, if a user A likes to watch action films and so does user B, then the movies that the user B will watch in the future will be recommended to A and vice-versa. Therefore, recommending movies is dependent on creating a relationship of similarities.

```{r}
similarity_mat <- similarity(ratingMatrix[1:4, ],
                               method = "cosine",
                               which = "users")
as.matrix(similarity_mat)

image(as.matrix(similarity_mat), main = "User's Similarities")
```

In the above matrix, each row and column represents a user. We have taken four users and each cell in this matrix represents the similarity that is shared between the two users.

Now, we delineate the similarity that is shared between the films.

```{r}
movie_similarity <- similarity(ratingMatrix[, 1:4], method =
                                 "cosine", which = "items")
as.matrix(movie_similarity)

image(as.matrix(movie_similarity), main = "Movies similarity")
```
Let us now extract the most unique ratings based on the movies.

```{r}
rating_values <- as.vector(ratingMatrix@data)
unique(rating_values) # extracting unique ratings
```
Now, we will create a table of ratings that will display the most unique ratings.

```{r}
Table_of_Ratings <- table(rating_values) # creating a count of movie ratings
Table_of_Ratings
```

#Provide explanation with any visuals.

In this step of the machine learning project, we will explore the most viewed movies in our dataset. We will first count the number of views in a film and then organize them in a table that would group them in descending order.

```{r}
library(ggplot2)
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(movie_data,
                                         movie_data$movieId == table_views[index,1])$title)
}
table_views[1:6,]
```

```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  ggtitle("Total Views of the Top Films")
```



